### 2021-01-23 paper/arm.tex
Updates

@@

We use the uninterpreted symbols $\Dx{\aLoc}$.

@@

\begin{definition}[$\xCO$/$\xRASC$/$\xDGA$]
  \label{def:pomsets-down}
  Update Def.~\ref{def:pomsets-ra} to:
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land((\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}
\end{definition}

\subsection{Read-Read dependencies ($\xRRD$)}

We use the uninterpreted symbol $\RW$.

\begin{definition}[$\xRRD$]
  \label{def:pomsets-rr}
  Update Def.~\ref{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm$,
  \end{enumerate}  
\end{definition}

We repeat all of the rules for preconditions and transformers,
Only \ref{L5} and \ref{T3} are changed from Def.~\ref{def:pomsets-down}.
We repeat \ref{L4}, although it is unchanged from Def.~\ref{def:pomsets-trans}.

\begin{definition}[$\xCO$/$\xRASC$/$\xDGA$/$\xRRD$]
  \label{def:pomsets-down-rr}
  Update Def.~\ref{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}  
\end{definition}

[Control dependencies into reads as in MP with release on right and control
dependency on left.]

@@

\subsection{Proof Strategy}

Alternate characterization of ARM:
\cite[\textsection B2.3.6]{arm-reference-manual}
\cite{alglave-git-alternate}
\cite{armed-cats}

>>>>

We discuss \armeight{} using \emph{external global completion} (\EGC{})
\cite{alglave-git-alternate,armed-cats}, which is very close to our model;
see also \cite[\textsection B2.3.6]{arm-reference-manual}.

@@

\begin{example}
The following execution is allowed by \armeight{}, but disallowed by
\refdef{def:pomsets-ra}.  The coherence order between the writes can be
witnessed by a separate thread, which we have elided.
\begin{gather*}
  \PW{x}{2}\SEMI 
  \PR[\mRA]{x}{r}\SEMI
  \PW{y}{1} \PAR
  \PW{y}{2}\SEMI
  \PW[\mRA]{x}{1}
  \\
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DW{x}{2}}{}
      \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
      \event{c}{\DW{y}{1}}{right=of b}
      \event{d}{\DW{y}{2}}{right=2.5em of c}
      \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
      \rf{a}{b}
      \sync{b}{c}
      \wk{c}{d}
      \sync{d}{e}
      \wk[out=-165,in=-15]{e}{a}
    \end{tikzinline}}
\end{gather*}
Under \EGC{}, this is explained by dropping the order
$\DRP[\mRA]{x}{2}\xsync\DWP{y}{1}$, because $\DRP[\mRA]{x}{2}$ is fulfilled
by a relaxed write in the same thread.
\begin{gather*}
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DW{x}{2}}{}
      \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
      \event{c}{\DW{y}{1}}{right=of b}
      \event{d}{\DW{y}{2}}{right=2.5em of c}
      \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
      \rf{a}{b}
      %\sync{b}{c}
      \wk{c}{d}
      \sync{d}{e}
      \wk[out=-165,in=-15]{e}{a}
    \end{tikzinline}}  
\end{gather*}

More generally, this can be understood as a compiler optimization that
downgrades a read from $\mRA$ to $\mRLX$ when it read can be fulfilled by a
relaxed write in the same thread.
\end{example}

To model such \emph{downgraded acquires}, we use the uninterpreted symbols
$\Dx{\aLoc}$.

@@

\begin{definition}[$\xCO$/$\xRASC$/$\xDGA$]
  \label{def:pomsets-down}
  Update \refdef{def:pomsets-ra} to:
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land((\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}
\end{definition}

% The following execution is allowed both by \armeight{} and \tso{}, but
% disallowed by \refdef{def:pomsets-ra}.
% \begin{gather*}
%   \PW{x}{2}\SEMI 
%   \PR[\mRA]{x}{r}\SEMI
%   \PR{y}{s} \PAR
%   \PW{y}{2}\SEMI
%   \PW[\mRA]{x}{1}
%   \\
%   \hbox{\begin{tikzinline}[node distance=1.5em]
%       \event{a}{\DW{x}{2}}{}
%       \raevent{b}{\DR[\mRA]{x}{2}}{right=of a}
%       \event{c}{\DR{y}{0}}{right=of b}
%       \event{d}{\DW{y}{2}}{right=2.5em of c}
%       \raevent{e}{\DW[\mRA]{x}{1}}{right=of d}
%       \rf{a}{b}
%       \sync{b}{c}
%       \wk{c}{d}
%       \sync{d}{e}
%       \wk[out=-165,in=-15]{e}{a}
%     \end{tikzinline}}
% \end{gather*}

\begin{example}
One might worry that our model is too permissive for $\mSC$ access, but
\armeight{} allows some counterintuitive results for SC access.  In the
following execution we elide the initializing write $\DWP{y}{0}$.
\begin{gather*}
  \IF{\PR{x}{}}\THEN \PW{x}{2} \FI\SEMI
  \PR[\mSC]{x}{r}\SEMI
  \PR[\mSC]{y}{s} \PAR
  \PW[\mSC]{y}{2}\SEMI
  \PW[\mSC]{x}{1}
  \\
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{a}{\DR{x}{1}}{}
      \event{b}{\DW{x}{2}}{right=of a}
      \scevent{c}{\DR[\mSC]{x}{2}}{right=of b}
      \scevent{d}{\DR[\mSC]{y}{0}}{right=of c}
      \scevent{e}{\DW[\mSC]{y}{2}}{right=2.5em of d}
      \scevent{f}{\DW[\mSC]{x}{1}}{right=of e}
      \po{a}{b}
      \rf{b}{c}
      %\sync{c}{d}
      \sync{e}{f}
      \wk{d}{e}
      \rf[out=-165,in=-15]{f}{a}
    \end{tikzinline}}
\end{gather*}
Under \EGC{}, this is explained by dropping the order
$\DRP[\mSC]{x}{2}\xsync\DRP{y}{0}$, because $\DRP[\mSC]{x}{2}$ is fulfilled
by a relaxed write in the same thread.
\end{example}

\subsection{Removing Read-Read dependencies ($\xRRD$)}

\begin{example}
The following execution is allowed by \armeight{}, but disallowed by
\refdef{def:pomsets-ra}.  
\begin{gather*}
  \begin{gathered}
  x\GETS0\SEMI 
  x\GETS 1\SEMI y^\mRA\GETS1 \PAR r\GETS y\SEMI \IF{r}\THEN s\GETS x\FI
  \\[-.4ex]
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{wx0}{\DW{x}{0}}{}
      \event{wx1}{\DW{x}{1}}{right=of wx0}
      \raevent{wy1}{\DW[\mRA]{y}{1}}{right=of wx1}
      \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
      \event{rx0}{\DR{x}{0}}{right=of ry1}
      \sync{wx1}{wy1}
      \po{ry1}{rx0}
      \rf{wy1}{ry1}
      \wk[in=-15,out=-165]{rx0}{wx1}
      \wk{wx0}{wx1}
    \end{tikzinline}}
\end{gathered}
\end{gather*}
Under \EGC{}, this is explained by dropping the order
$\DRP{y}{1}\xpo\DRP{x}{0}$, because \armeight{} does not include control
dependencies between reads in the locally-ordered-before relation.
\begin{gather*}
  \hbox{\begin{tikzinline}[node distance=1.5em]
      \event{wx0}{\DW{x}{0}}{}
      \event{wx1}{\DW{x}{1}}{right=of wx0}
      \raevent{wy1}{\DWRel{y}{1}}{right=of wx1}
      \event{ry1}{\DR{y}{1}}{right=2.5em of wy1}
      \event{rx0}{\DR{x}{0}}{right=of ry1}
      \sync{wx1}{wy1}
      %\po{ry1}{rx0}
      \rf{wy1}{ry1}
      \wk[in=-15,out=-165]{rx0}{wx1}
      \wk{wx0}{wx1}
    \end{tikzinline}}  
\end{gather*}
\end{example}

Since we do not distinguish control dependencies from other dependencies, we
are forced to drop all dependencies between reads.  In order to do so, we use
the uninterpreted symbol $\RW$.

\begin{definition}[$\xRRD$]
  \label{def:pomsets-rr}
  Update \refdef{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm$,
  \end{enumerate}  
\end{definition}

We repeat all of the rules for preconditions and transformers,
Only \ref{L5} and \ref{T3} are changed from \refdef{def:pomsets-down}.
We repeat \ref{L4}, although it is unchanged from \refdef{def:pomsets-trans}.

\begin{definition}[$\xCO$/$\xRASC$/$\xDGA$/$\xRRD$]
  \label{def:pomsets-down-rr}
  Update \refdef{def:pomsets-trans} to:
  \begin{enumerate}
  \item[\ref{S3})]
    $\labelingForm(\aEv)$ implies $\QS{\aLoc}{\amode}\land\aExp{=}\aVal$,
  \item[\ref{L3})]
    $\labelingForm(\aEv)$ implies $\QL{\aLoc}{\amode}$.
  \item[\ref{T3})]
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\TRUE/\RW]$ if $\labelingAct_1(\aEv)$ is a write,
    \\
    $\labelingForm(\aEv)$ implies
    $\labelingForm_1(\aEv)[\TRUE/\Q{}][\FALSE/\RW]$ otherwise.
  \end{enumerate}
  \begin{enumerate}
  \item[\ref{S4})]
    $\aTr{\bEvs}{\bForm}$ implies $(\Qw{\aLoc}\limplies\aExp{\neq}\aVal)\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$,
  \item[\ref{S5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qw{\aLoc}\land \DS{\aLoc}{\amode}\bForm[\aExp/\aLoc]$.
  \item[\ref{L4})]
    $\aTr{\bEvs}{\bForm}$ implies $\aVal{=}\aReg\limplies\bForm$, 
  \item[\ref{L5})]
    $\aTr{\cEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land(\RW\limplies(\aVal{=}\aReg\lor\aLoc{=}\aReg)\limplies\bForm)$,
  \item[\ref{L6})]
    $\aTr{\dEvs}{\bForm}$ implies $\lnot\Qx{\aLoc} \land \DL{\aLoc}{\amode} \land\bForm$.
  \end{enumerate}  
\end{definition}

@@

%\subsection{Proof Strategy}
